### 2022-11-14 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
fixing path changes exploratory

---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output: html_document
---