### 2022-11-17 Code/GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
Did as much as I could on survey data

---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output: html_document
---

## Introduction

In the great pursuit of mitigating climate change, countries have differed greatly in their strategies and successes. Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D. C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.] As other countries trail behind, illuminating how Denmark is successfully reducing its emissions is pertinent for other countries. Although reducing emissions involves extensive effort from multiple sectors, the citizens also greatly determine the success of mitigation plans.^[Weber, E. U., "What Shapes perceptions of climate change?," WIRES Climate Change.2010] What determines people's perceptions and willingness to decrease their own emissions, is therefore important for countries to understand. To delve into this question data from two global surveys on climate perceptions will be analyzed along with green house gas emissions and the reduction rate of countries. Special emphasis will be on comparing Denmark due to the countrie's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings. 2022.]

This data analysis will consist of three sections: the greenhouse gas emissions of the U.S. and Denmark, the climate change perceptions of their citizens, and the correlation between emissions and climate change perceptions, using data from many countries.

In the emissions section Denmark's the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]

The second section will include an analysis of several questions regarding climate change for both the U.S. and Denmark. These will include what Americans and Danes believe causeds climate change, how worried they are about it, and how big of a threat they believe climate change is in the next 20 years. We hypothesize Denmark will have a significantly higher percentage of people who are believe climate change is human caused than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane. “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502. DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more Danes will be very worried about climate change and see it as an extreme threat than U. S. citizens, because climate change is uniformly depicted in Danish media as a problem, but it is not in the U.S.^4

In the third section we will overlay the survey data with the emissions data to determine the percentage of causation. We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang. C., " The influence of political ideology on greenhouse gas emissions" Global Environmental Change 74, no.1 (May 2022):102496. DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]

## Statistcal Analysis and Results

Statistical analysis will consist of the aforementioned three sections.

## Climate Change Emissions of Countries

Preparing R
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
library(forecast)
```

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.

Lets first view the emissions of Denmark and the United States.

Reading in the Data
```{r}
emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"))
```

Lets turn the date column into date data using the lubridate package.  
To do that we need to put in the day and the month using string.

```{r}
emissions$year<-paste0("01-01-", emissions$year)
```

???This NA'd all my data and %y messed my dates up so I had to do the code below???
```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

#Now when running statistical tests R will know the data are dates.

### Creating the figure

???HOW DO I GET RID OF THE LINE?
```{r}

#year <- function(x) as.POSIXlt(x)$year + 1900
# I do not know how the above code does this :(
ggplot(emissions, aes(year, emissions)) + 
  geom_path() + geom_point(aes(colour=country))+ theme_bw()

```
Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States.

### Running the Test

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```

Turning our data into a time series

```{r}
emissions_ts<-ts(emissions$emissions, start=c(1990,01), end=c(2019,01), frequency=1)
```

```{r}
head(emissions_ts)
plot(emissions_ts, xlab="year", ylab="emissions")
```
Making a moving average:
```{r}
Ma_emissions<- forecast::ma(emissions_ts, order=3, centre = TRUE)
```

```{r}
plot(emissions_ts)
lines(Ma_emissions, col="maroon", lwd=3)
```
Not sure if I am supposed to do this next part...

```{r}
seasonal_add<-emissions_ts - Ma_emissions
plot(seasonal_add)
```
ya.. this looks weird.

```{r}
fit_add<-decompose(emissions_ts, type='additive')
plot(fit_add)
```
Ya, so I cannot decompose my data. She also did not teach us how to run the test, which is annoying.

### Rate of Emissions

To determine rate the emissions must be divided by the year.

???? HOW DO THAT WITH TIME SERIES???

## Perceptions of Climate Change

Reading in the data:
```{r}
survey<- read.csv(here("Data for the Code", "Corrected_Data_sets", "final_ccos_US_Denmark.csv"), stringsAsFactors=TRUE )
```

Though more people may have been interviewed in one country than the other, the survey data is in percent, as in the percentage of the whole who chose that option. We can double check to ensure the recorded numbers equals 100 per country per survey question.

Lets check with the 'beilefs question' which is the first we will anaylzed.

```{r}
#Parcing out just the survey answers regarding climate change beliefs
beliefs<- survey %>% filter(survey$question=="Ques__Climate_beliefs") 
```

```{r}
total<- beliefs %>% group_by(Country) %>% summarise(total=sum(Number))
```

This means we can fairly compare Danes and American's perceptions of climate change.

### Beliefs of Climate Change

Lets first see what Danes and Americans believe caused climate. The options are the survey are: caused mostly by humans, caused by natural change in the environment, caused by both, and climate change is not happening.

```{r}
# Making the graph
ggplot(beliefs, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  xlab("Response to Survey Question")+ ylab("Percent") + # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```
More Americans believe that climate change is not happening or that it is caused by natural changes than Danes. More Danes believe it is caused by human activities. However, more Danes than Americans think it is cause equally by both. Lets run a statistical test to see if Danes believe climate change is human caused significantly more than Americans.

```{r}
ggplot(beliefs, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```
The pie chart clearly shows that most of the Danes surveyed believe climate change is human caused, whereas more Americans thought it was a mix of human and natural causes or that climate change is not happening.

We have multiple categorical predictor variables with the response variable being ratio data that bounded between 0 and 100%, so we will run a generalized linear model.

### Running the Statistical Test

The family is not binomial, poisson, or gamma, ugh what is it? It accepted gaussain but autoplot looked wrong.

Gaussian family : for continuous decimal data with normal distribution, like weight, length, et al. frequency. So given I have percentage data, which is similar, I believe I should use poisson. However,

```{r}
belief_glm<-glm(Number~response, data=beliefs, family= Gamma(link='inverse') ) #Constructing the general linear model
#The family (or ‘error’) part of a GLM is not too difficult to understand. This bit just determines what kind of distribution is used to describe the response variable. A general linear model always assumes normality— that’s pretty much its defining feature in fact—but with a generalized linear

#changing the default link yoou doo family=poiossoono(link=log)) for log. Remember, this is the log link function for Poisson models. If you want to be explicit about the link function, or change the default, this

```
Test autoplot

```{r}
autoplot(belief_glm)
```
Something is still wrong

```{r}
anova(belief_glm)
```
???I believe I am not choosing the right family=.??? 
???? Or I have really underdispersed data?

### Level of worriedness about Climate Changed
In this survey question people were asked how worried they are about climate change. There are three options for 
```{r}
worry<- survey %>% filter(survey$question=="Ques__climate_worry") 
```

```{r}
ggplot(worry, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
```
32% of Americans are very worried about climate change, wherease oonly 21% of Danes are. But 52% of Danes are somewhat worried at only 36% are. To better visualize this lets try another graph.

```{r}
ggplot(worry, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```
We can see that most Danes are somewhat worried, with the smallest percentage being not worried at all. Most Americans are also somewhat worried, though the percentage is smaller compared too Denmark (51% to 35.6%). More Americans however are very worried, and more Americans are not at all worried than Danes. Interesting. A statistical test is needed to see if Americans are less worried than Danes.

### Lets do a statistical test to verify. 
Again we will do a general linear model.

```{r}
worry_glm<-glm(Number~response, data=worry, family= Gamma(link='inverse') ) #Constructing the

```
Test autoplot

```{r}
autoplot(worry_glm)
```
again, very concerning

```{r}
anova(worry_glm)
```
## Is Climate Change a Big Threat

In this survey question people were asked whether or not climate change is a big threat within the next 20 years. The options were very serious threat, somewhat serious, not a threat at all, or I do not know.

parcing the data
```{r}
threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
```

```{r}
ggplot(threat, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + xlab("Response to Survey Question")+ ylab("Percent") + theme_bw() + coord_flip()
```
More Americans than Danes think that climate change is a very serious threat, but Danes think it is a somewhat serious threat than Americans. More Americans also think that cliamte change is not a threat at all than Danes. Lets view a pie chart.

```{r}
ggplot(threat, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```
We can see that most Danes, %53 think climate change is a somewhat serious threat, wherease most Americans, %40.6, think it is a very serioous threat. Only 17.5% of Danes think that it is not a threat at all, compared to the 22.3% of Americans that think that. Lets test with a statistical test.

### Lets do a statistical test to verify. 
Again we will do a general linear model.

```{r}
threat_glm<-glm(Number~response, data=threat, family= Gamma(link='inverse') ) #Constructing the

```
Test autoplot

```{r}
autoplot(threat_glm)
```
again, very concerning

```{r}
anova(threat_glm)
```
## More or less Fossil Fuels

Now that we know what Danes and Americans think about climate change and how worried they are, lets see how they are reacting. This next survey question asked people if their region should use more or less fossil fuels with the options being much more, somewhat more, same as today (same amount), somewhat less, or much less.

Lets check with the 'beilefs question' which is the first we will anaylzed.

```{r}
#Parcing out just the survey answers regarding climate change beliefs
fuels<- survey %>% filter(survey$question=="Ques__Fossil_moreorless") 
```

```{r}
# Making the graph
ggplot(fuels, aes(response,Number, fill=Country))+geom_bar(stat = 'identity', position = 'dodge') + 
  xlab("Response to Survey Question")+ ylab("Percent") + # renaming x and y axis labels
  theme_bw() +
  coord_flip() # switching x and y axis to read the suvey questions
```
More Americans think there should be much more or or somewhat more fossil fuels used than Danes. More Danes thinks threre should be somewhat less or much less fossil fuels used. This appears to clearly show that Danes think fossil fuel usage should be decreased, where at Americans think that less. Lets see in pie char form.

```{r}
ggplot(fuels, aes(x="", y= Number, fill=response))+  geom_bar(stat="identity", width=1)+ coord_polar("y", start=0) + 
   geom_label(aes(label = round(Number, digits = 1)), nudge_x = 0.4) + # add percentages
  theme_void()+ # remove background, grid, numeric labels
facet_wrap(~Country)
```
The pie chart clearly shows that most of the Danes surveyed, %53.6, believed much less fossil fuels should be used, with the second largest percentage of 25.7% believed somewhat less should be used. These percentages are larger than results in the United states, though the largest percentage of Americans also believed there should be much less fossil fuel usage at 40.5%, with the second most thinking somewhat less fossil fuel usage, 17.9%. More Americans believed there should be much more or somewhat more than Danes. Lets verify with a statistical test.

we will run another generalized linear model.

### Running the Statistical Test

```{r}
fuels_glm<-glm(Number~response, data=fuels, family= Gamma(link='inverse') ) #Constructing the 
```
Test autoplot

```{r}
autoplot(fuels_glm)
```
Something is still wrong

```{r}
anova(fuels_glm)
```
???I believe I am not choosing the right family=.??? 
???? Or I have really underdispersed data?

## Survey data Impacting Emissions

## Summary Discussion

## Challenges and Limitations